---
title: "PollutionExploration"
author: "James Budarz"
date: "July 16, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(lubridate)
```

# Questions worth asking:
0. For splitting data into danger vs safe zone: What are high vs. safe levels? Asthma etc. <https://www.epa.gov/criteria-air-pollutants/naaqs-table>

1. Which cities are most polluted? The least? How does city population affect pollution levels?

2. How does the pollution change over the long term?

3. How does the pollution change over the short term?

4. Are there any remarkable events?

5. Plot values on the map

```{r loaddata, include=FALSE, cache=TRUE}
df = read.csv("data/uspollution/pollution_us_2000_2016.csv", header = TRUE)
```

##Step 0: For exploratory purposes, just deal with Arizona data:

```{r Data Cleanup, include=FALSE}
azdata = df %>% filter(State == "Arizona")
```

##Step 1: Clean the data (remove locations that don't have daily data)
```{r Data Reduction}
# Find number of measurements for each location:
azdata %>% group_by(Site.Num) %>%
  summarize(first(City), first(Address),NumberOfMeasurements = n())

# Convert measurements to a common unit:

```

##Step 2: Refine the data
```{r fetchlocation}
# Fetch location (lat, long) information for every address

# Assign each date a month
azdata$Date.Local = as.Date(azdata$Date.Local)
azdata <- azdata %>%
  mutate(month = factor(
    months(Date.Local), levels = c('January','February','March','April','May','June',
                                   'July','August','September','October','November','December')
    ))
# Remove strange data before 2001:
azdata <- azdata %>% filter(year(Date.Local) > 2001)

```

##Step 3: Make some plots of interest:
```{r Data Plotting}
gg <- ggplot(azdata)

gg + geom_point(aes(x = CO.Mean, y = NO2.Mean, color = City), size = .05) +
  ggtitle('NO2 Pollution Levels in Arizona')

gg + geom_point(aes(x = Date.Local, y = NO2.Mean, color = City), size = .05) +
  ggtitle('NO2 Pollution Levels in Arizona') +
  geom_smooth(aes(x = Date.Local, y = NO2.Mean, color = City), formula = y ~ sin(x) + x)

gg + geom_col(aes(x = month, y = NO2.Mean, fill = City), position = 'dodge') +
  ggtitle('NO2 Levels in AZ by Month')

```


